---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ERDbuilder

<!-- badges: start -->

<!-- badges: end -->

This R package creates Entity-Relationship Diagrams (ERDs). It generates traditional ERDs from data frames, enabling users to specify relationship cardinalities and variables involved in joins. The package also facilitates data joins based on the established ERD.

## Package Dependencies

-   `DiagrammeR`

-   `dplyr`

## Objective

Create Entity Relationship Diagrams (ERD) by extracting the attribute names of tables to create complex graphs where:

-   Nodes: Each entity (i.e., data frame) is represented as a node. The node label consists of the entity name and the attribute names within the entity.
-   Edges: Relationships between entities are represented as edges between the corresponding nodes. Labels at the edges' ends indicate the relationship's type and cardinality.

## Installation

You can install the development version of ERDbuilder like so:

``` r
remotes::install_github("gbasulto/ERDbuilder")
```

## Function and Object List

| Name             | Class    | Purpose                     |
|------------------|----------|-----------------------------|
| `render_erd()`   | Function | Render ERD                  |
| `perform_join()` | Function | Perform left join of tables |
| `create_erd()`   | Function | Create ERD object           |

: Functions and Objects in "ERDbuilder" package

## Example 1

This is a basic example which shows you how to solve a common problem:

```{r example2, eval = TRUE}

## Load packages _______________________________________________________________
library(ERDbuilder)

## Create datasets _____________________________________________________________

# Create dataframe "employees"
employees <- data.frame(
  emp_id = c(1, 2, 3),
  name = c("Alice", "Bob", "Charlie")
)

# Create dataframe "departments"
departments <- data.frame(
  dept_id = c(1, 2),
  dept_name = c("HR", "Engineering")
)

# Create dataframe "assignments"
assignments <- data.frame(
  emp_id = c(1, 3),
  dept_id = c(1, 2)
)

# Define relationships _________________________________________________________

relationships <- list(
  assignments = list(
    employees = list(emp_id = "emp_id", relationship = c("||", "||")),
    departments = list(dept_id = "dept_id", relationship = c(">0", "||"))
  )
)

# Create ERD object ____________________________________________________________
erd_object <- create_erd(
  list(
    employees = employees, 
    departments = departments, 
    assignments = assignments), 
  relationships)

# Render the ERD graphically ___________________________________________________
edr_plot2 <- render_erd(erd_object, label_distance = 0, label_angle = -25)

# Plot ERD _____________________________________________________________________
edr_plot2

# ## Un-comment to export to TIFF ______________________________________________
# DPI = 600
# WidthCM = 38
# HeightCM = 38
# 
# edr_plot2 |> 
#   export_svg() |> 
#   charToRaw |> 
#   rsvg(width = WidthCM * (DPI / 2.54), 
#        height = HeightCM *(DPI / 2.54)) |> 
#   tiff::writeTIFF("edr_plot2.tiff")

```

## Example 2

```{r example1, eval = TRUE, warning=FALSE, message=FALSE}

## Load packages ______________________________________________________________
library(ERDbuilder)
library(dplyr)
library(readr)

## Set URLs ____________________________________________________________________
data_url <- "https://raw.githubusercontent.com/jwood-iastate/DataFiles/main/"
occ_url <- paste0(data_url, "OCC.csv")
crash_url <- paste0(data_url, "CRASH.csv")
distract_url <- paste0(data_url, "DISTRACT.csv")
vehicle_url <- paste0(data_url, "GV.csv")
  
## Load data ___________________________________________________________________
occupant_tbl <- read_csv(occ_url, show_col_types = FALSE)     # Occupant Data
crash_tbl <- read_csv(crash_url, show_col_types = FALSE)      # Crash data
distract_tbl <- read_csv(distract_url, show_col_types = FALSE)# Distraction data
vehicle_tbl <- read_csv(vehicle_url, show_col_types = FALSE)  # Vehicle data

## Define relationships ________________________________________________________

relationships <- list(
  Crash = list(
    Vehicle = list(
      CASENUMBER = "CASENUMBER", relationship = c("||", "|<")),
    Occupant = list(
      CASENUMBER = "CASENUMBER", relationship = c("||", "|<")),
    Distract = list(
      CASENUMBER = "CASENUMBER", relationship = c("||", "0<"))
  ),
  Vehicle = list(
   # Crash = list(
   #   CASENUMBER = "CASENUMBER", relationship = c("|<", "||")),
    Occupant = list(
      CASENUMBER = "CASENUMBER", VEHNO = "VEHNO", relationship = c("|0", "0<")),
    Distract = list(
      CASENUMBER = "CASENUMBER", VEHNO = "VEHNO", relationship = c("||", "0<"))
  )
)

## Create the ERD object _______________________________________________________

erd_object <- 
  create_erd(
    list(
      Crash = crash_tbl, 
      Vehicle = vehicle_tbl, 
      Occupant = occupant_tbl, 
      Distract= distract_tbl), 
    relationships)

# Perform joins ________________________________________________________________ 

# Note that there will be a many-to-many relationship when joining the Distract
# table since the Crash, Vehicle, and Occupant tables will have already been
# joined.
joined_data <- 
  perform_join(erd_object, c("Crash", "Vehicle", "Occupant", "Distract"))

## Render plot _________________________________________________________________

edr_plot <- 
  render_erd(erd_object, label_distance = 0, label_angle = 15, n = 20)

edr_plot

## Uncomment to export to SVG __________________________________________________

# # Render the ERD graphically, save as a .tiff, then include in rendered files
# library(rsvg)
# library(DiagrammeRsvg)
# 
# 
# DPI <- 600
# WidthCM <- 38
# HeightCM <- 38
# 
# edr_plot |> 
#   export_svg() |> 
#   charToRaw() |> 
#   rsvg(
#     width = WidthCM * (DPI / 2.54), 
#     height = HeightCM * (DPI / 2.54)) |> 
#   tiff::writeTIFF("edr_plot.tiff")

```
